Perturbing SU($n$)-symmetric interactions

### 2020-02-05 collective_sun/sun_perturbation.tex
simplified diagrams for triple product of three two-body operators

Calculating the second-order effective Hamiltonian $H_1^{(2)}$ induced
on the permutationally symmetric manifold $\M_0$ by the perturbation
$\V_1$ in \eqref{eq:perturbations} requires us to simplify a product
of the form
\begin{align}
  \P_0 \sp{\sum_{p,q} v_p w_q X_p Y_q} \P_0
  \EQFS \diagram{single_body_2} X_1 Y_2
  + \diagram{single_body_1} X_1 Y_1,
  \label{eq:PXYP_start}
\end{align}
where $p,q$ index individual spins; $X,Y$ are single-spin operators;
and the coefficients $v_p,w_q$ satisfy \eqref{eq:cond_1} with
eigenvalue $\Delta\ne0$, which implies that they are mean-zero.
Defining $\v m \equiv \sum_p m_p \ket{p}$ for $m\in\set{v,w}$, we can
simplify
\begin{align}
  \diagram{single_body_1}
  \equiv \sum_p v_p w_p
  = \v v\c\v w,
\end{align}
and
\begin{align}
  \diagram{single_body_2}
  = \diagram{single_body_2_o} - \diagram{single_body_2_x}
  = \diagram{single_body_2_oo} - \diagram{single_body_1}
  = - \v v \c\v w,
\end{align}
so
\begin{align}
  \sum_{p,q} v_p w_q X_p Y_q
  \EQFS - \v v\c\v w \p{X_1 Y_2 - X_1 Y_1},
\end{align}
In order to write this result in terms of collective operators
$\col{Z} \equiv \sum_p Z_p$, we expand
\begin{align}
  \col{X Y} \EQFS N X_1 Y_1,
  &&
  \col{X}\,\col{Y}
  = \sum_{p=q} X_p Y_q + \sum_{p\ne q} X_p Y_q
  \EQFS \col{XY} + N\p{N-1} X_1 Y_2,
\end{align}
which implies that
\begin{align}
  \sum_{p,q} v_p w_q X_p Y_q
  \EQFS - \f{\v v \c\v w}{N\p{N-1}}
  \p{\col{X}\,\col{Y} - N \col{XY}}.
\end{align}

@@

Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
on the permutationally symmetric manifold $\M_0$ by the perturbation
$\V_2$ in \eqref{eq:perturbations} requires us to simplify a product
of the form
\begin{multline}
  \P_0 \sp{\sum_{k<\ell} \sum_{p<q}
    v_{k\ell} w_{pq} O_{k\ell} Q_{pq}} \P_0 \\
  \EQFS \f14 \diagram{two_body_4} O_{1,2} Q_{3,4}
  + \diagram{two_body_3} O_{1,2} Q_{1,3}
  + \f12 \diagram{two_body_2} O_{1,2} Q_{1,2},
  \label{eq:POQP_start}
\end{multline}
where $k,\ell,p,q$ index individual spins; $O$ and $Q$ are
permutationally-symmetric two-spin operators; the coefficients
$v_{k\ell},w_{pq}$ satisfy \eqref{eq:cond_2} with eigenvalue
$\Delta\ne0$, which implies that they are mean-zero; and the factors
of $1/4$ and $1/2$ are included to cancel out the symmetry factors
that appear in the definition of a diagram.  Defining
\begin{align}
  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
  &&
  \v m \equiv \sum_p m_p \ket{p},
  &&
  m_p \equiv \sum_q m_{pq},
\end{align}
for $m\in\set{v,w}$, we can write
\begin{align}
  \diagram{two_body_2} = \sum_{p,q\in\ZZ_N} v_{pq} w_{pq}
  = 2 \sum_{p,q\in\C_N\p{2}} v_{pq} w_{pq}
  = \v{\m v} \c\v{\m w},
\end{align}
and simplify
\begin{align}
  \diagram{two_body_3}
  &= \diagram{two_body_3_o} - \diagram{two_body_3_x} \\
  &= \diagram{two_body_3_oo} - \diagram{two_body_3_ox}
  - \diagram{two_body_2} \label{eq:mid_zero} \\
  &= \v v \c \v w - 2 \v{\m v} \c \v{\m w},
\end{align}
where the middle diagram in \eqref{eq:mid_zero} contains an empty sum
for the cross, and is therefore equal to zero.  Similarly neglecting
diagrams with empty sums and additionally using the fact that
$v_{pq},w_{pq}$ are mean-zero, we can work out
\begin{align}
  \diagram{two_body_4}
  &= \diagram{two_body_4_o} - \diagram{two_body_4_x} \\
  &= \sp{\diagram{two_body_4_oo} - \diagram{two_body_4_ox}}
  - 2 \sp{\diagram{two_body_3}} \\
  &= \sp{-2 \diagram{two_body_3_o}}
  - 2 \sp{\diagram{two_body_3_oo} - \diagram{two_body_2}} \\
  &= -4 \diagram{two_body_3_oo} + 2 \diagram{two_body_2} \\
  &= -4 \v v\c\v w + 4 \v{\m w}\c\v{\m w}.
\end{align}
Altogether, we thus have that
\begin{multline}
  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} O_{k\ell} Q_{pq} \\
  \EQFS - \p{\v v \c \v w - \v{\m v} \c \v{\m w}} O_{1,2} Q_{3,4}
  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} O_{1,2} Q_{1,3}
  + \v{\m v} \c \v{\m w}\, O_{1,2} Q_{1,2}.
  \label{eq:POQP}
\end{multline}

@@

Here we simplify a product of three two-body operators of a particular
form projected onto the permutationally symmetric manifold:
\begin{align}
  \X \equiv \P_0 \sp{\sum_{k<\ell} \sum_{p<q} \sum_{r<s}
    u_{k\ell} v_{pq} w_{rs} X_k X_\ell X_p X_q X_r X_s} \P_0,
  &&
  X^2 = 1,
\end{align}
where $k,\ell,p,q,r,s$ index individual spins; $X$ is a single-spin
operator; and $u_{k\ell},v_{pq},w_{rs}$ are arbitrary scalar
coefficients with $m_{pq}=m_{qp}$ and $m_{pp}=0$ for each of
$m\in\set{u,v,w}$.  Collecting terms according to the number of $X$
operators that remain after accounting for $X^2=1$, we find that
\begin{align}
  \X \EQFS A_6 X^{\otimes 6} + A_4 X^{\otimes 4} + A_2 X^{\otimes 2},
  \label{eq:triple_multi}
\end{align}
with
\begin{align}
  A_6 \equiv \f18 \diagram{triple_6},
  &&
  A_4 \equiv \diagram{triple_4_1} + \f32 \diagram{triple_4_0},
  \label{eq:triple_64}
\end{align}
\begin{align}
  A_2 \equiv \f12 \diagram{triple_2_0} + 3 \diagram{triple_2_1}
  + \f34 \diagram{triple_2_2} + 3 \diagram{triple_2_3},
  \label{eq:triple_2}
\end{align}
where a diagram with unlabeled circles denotes an average over all
label assignments that result in a distinct diagram, e.g.
\begin{align}
  \diagram{triple_6} \equiv \diagram{triple_6_uvw},
  \label{eq:triple_uvw}
\end{align}
\begin{align}
  \diagram{triple_2_3}
  \equiv \f13 \sp{\diagram{triple_2_3_uvw}
  + \diagram{triple_2_3_vwu} + \diagram{triple_2_3_wuv}}.
  \label{eq:triple_uvw_3}
\end{align}
The first diagram in each of $A_6,A_4,A_2$ as written in
\eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment
of labels, as in \eqref{eq:triple_uvw}, while all other unlabeled
diagrams have three assignments, as in \eqref{eq:triple_uvw_3}.  We
can write \eqref{eq:triple_multi} in terms of collective operators as
\begin{align}
  \X \EQFS
  \tilde A_6 \col{X}^6 + \tilde A_4 \col{X}^4 + \tilde A_2 \col{X}^2,
  \label{eq:triple_col}
\end{align}
where we have neglected scalar terms without physical consequence, and
\begin{align}
  \tilde A_6 \equiv \f{A_6}{N!_6},
  &&
  \tilde A_4 \equiv \f{A_4}{N!_4} - 5\p{3N-8} \f{A_6}{N!_6},
\end{align}
\begin{align}
  \tilde A_2 \equiv \f{A_2}{N!_2} - 2\p{3N-4} \f{A_4}{N!_4}
  + \p{45N^2-210N+184} \f{A_6}{N!_6}.
\end{align}

>>>>

Calculating the second-order effective Hamiltonian $H_1^{(2)}$ induced
on the permutationally symmetric manifold $\M_0$ by the perturbation
$\V_1$ in \eqref{eq:perturbations} requires us to simplify a product
of the form
\begin{align}
  \P_0 \sp{\sum_{p,q} v_p w_q X_p Y_q} \P_0
  \EQFS \diagram{single_body_0} X_1 Y_2
  + \diagram{single_body_1} X_1 Y_1,
  \label{eq:PXYP_start}
\end{align}
where $p,q$ index individual spins; $X,Y$ are single-spin operators;
and the coefficients $v_p,w_q$ satisfy \eqref{eq:cond_1} with
eigenvalue $\Delta\ne0$, which implies that they are mean-zero.
Defining $\v m \equiv \sum_p m_p \ket{p}$ for $m\in\set{v,w}$, we can
simplify
\begin{align}
  \diagram{single_body_1}
  \equiv \sum_p v_p w_p
  = \v v\c\v w,
\end{align}
and
\begin{align}
  \diagram{single_body_0}
  = \diagram{single_body_0_o} - \diagram{single_body_0_x}
  = \diagram{single_body_0_oo} - \diagram{single_body_1}
  = - \v v \c\v w,
\end{align}
so
\begin{align}
  \sum_{p,q} v_p w_q X_p Y_q
  \EQFS - \v v\c\v w \p{X_1 Y_2 - X_1 Y_1},
\end{align}
In order to write this result in terms of collective operators
$\col{Z} \equiv \sum_p Z_p$, we expand
\begin{align}
  \col{X Y} \EQFS N X_1 Y_1,
  &&
  \col{X}\,\col{Y}
  = \sum_{p=q} X_p Y_q + \sum_{p\ne q} X_p Y_q
  \EQFS \col{XY} + N\p{N-1} X_1 Y_2,
\end{align}
which implies that
\begin{align}
  \sum_{p,q} v_p w_q X_p Y_q
  \EQFS - \f{\v v \c\v w}{N\p{N-1}}
  \p{\col{X}\,\col{Y} - N \col{XY}}.
\end{align}

@@

Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
on the permutationally symmetric manifold $\M_0$ by the perturbation
$\V_2$ in \eqref{eq:perturbations} requires us to simplify a product
of the form
\begin{multline}
  \P_0 \sp{\sum_{k<\ell} \sum_{p<q}
    v_{k\ell} w_{pq} O_{k\ell} Q_{pq}} \P_0 \\
  \EQFS \f14 \diagram{two_body_0} O_{1,2} Q_{3,4}
  + \diagram{two_body_1} O_{1,2} Q_{1,3}
  + \f12 \diagram{two_body_2} O_{1,2} Q_{1,2},
  \label{eq:POQP_start}
\end{multline}
where $k,\ell,p,q$ index individual spins; $O$ and $Q$ are
permutationally-symmetric two-spin operators; the coefficients
$v_{k\ell},w_{pq}$ satisfy \eqref{eq:cond_2} with eigenvalue
$\Delta\ne0$, which implies that they are mean-zero; and the factors
of $1/4$ and $1/2$ are included to cancel out the symmetry factors
that appear in the definition of a diagram.  Defining
\begin{align}
  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
  &&
  \v m \equiv \sum_p m_p \ket{p},
  &&
  m_p \equiv \sum_q m_{pq},
\end{align}
for $m\in\set{v,w}$, we can write
\begin{align}
  \diagram{two_body_2} = \sum_{p,q\in\ZZ_N} v_{pq} w_{pq}
  = 2 \sum_{p,q\in\C_N\p{2}} v_{pq} w_{pq}
  = \v{\m v} \c\v{\m w},
\end{align}
and simplify
\begin{align}
  \diagram{two_body_1}
  &= \diagram{two_body_1_o} - \diagram{two_body_1_x} \\
  &= \diagram{two_body_1_oo} - \diagram{two_body_1_ox}
  - \diagram{two_body_2} \label{eq:mid_zero} \\
  &= \v v \c \v w - 2 \v{\m v} \c \v{\m w},
\end{align}
where the middle diagram in \eqref{eq:mid_zero} contains an empty sum
for the cross, and is therefore equal to zero.  Similarly neglecting
diagrams with empty sums and additionally using the fact that
$v_{pq},w_{pq}$ are mean-zero, we can work out
\begin{align}
  \diagram{two_body_0}
  &= \diagram{two_body_0_o} - \diagram{two_body_0_x} \\
  &= \sp{\diagram{two_body_0_oo} - \diagram{two_body_0_ox}}
  - 2 \sp{\diagram{two_body_1}} \\
  &= \sp{-2 \diagram{two_body_1_o}}
  - 2 \sp{\diagram{two_body_1_oo} - \diagram{two_body_2}} \\
  &= -4 \diagram{two_body_1_oo} + 2 \diagram{two_body_2} \\
  &= -4 \v v\c\v w + 4 \v{\m w}\c\v{\m w}.
\end{align}
Altogether, we thus have that
\begin{multline}
  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} O_{k\ell} Q_{pq} \\
  \EQFS - \p{\v v \c \v w - \v{\m v} \c \v{\m w}} O_{1,2} Q_{3,4}
  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} O_{1,2} Q_{1,3}
  + \v{\m v} \c \v{\m w}\, O_{1,2} Q_{1,2}.
  \label{eq:POQP}
\end{multline}

@@

Here we simplify a product of three two-body operators of a particular
form projected onto the permutationally symmetric manifold:
\begin{align}
  \X \equiv \P_0 \sp{\sum_{k<\ell} \sum_{p<q} \sum_{r<s}
    u_{k\ell} v_{pq} w_{rs} X_k X_\ell X_p X_q X_r X_s} \P_0,
  &&
  X^2 = 1,
\end{align}
where $k,\ell,p,q,r,s$ index individual spins; $X$ is a single-spin
operator; and $u_{k\ell},v_{pq},w_{rs}$ are arbitrary scalar
coefficients with $m_{pq}=m_{qp}$ and $m_{pp}=0$ for each of
$m\in\set{u,v,w}$.  Collecting terms according to the number of $X$
operators that remain after accounting for $X^2=1$, we find that
\begin{align}
  \X \EQFS A_6 X^{\otimes 6} + A_4 X^{\otimes 4} + A_2 X^{\otimes 2},
  \label{eq:triple_multi}
\end{align}
with
\begin{align}
  A_6 \equiv \f18 \diagram{triple_0},
  &&
  A_4 \equiv \diagram{triple_1} + \f32 \diagram{triple_01},
  \label{eq:triple_64}
\end{align}
\begin{align}
  A_2 \equiv \f12 \diagram{triple_2} + 3 \diagram{triple_11}
  + \f34 \diagram{triple_02} + 3 \diagram{triple_011},
  \label{eq:triple_2}
\end{align}
where a diagram with unlabeled circles denotes an average over all
label assignments that result in a distinct diagram, e.g.
\begin{align}
  \diagram{triple_0} \equiv \diagram{triple_0_uvw},
  \label{eq:triple_uvw}
\end{align}
\begin{align}
  \diagram{triple_011}
  \equiv \f13 \sp{\diagram{triple_011_uvw}
  + \diagram{triple_011_vwu} + \diagram{triple_011_wuv}}.
  \label{eq:triple_uvw_3}
\end{align}
The first diagram in each of $A_6,A_4,A_2$ as written in
\eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment
of labels, as in \eqref{eq:triple_uvw}, while all other unlabeled
diagrams have three assignments, as in \eqref{eq:triple_uvw_3}.

Having identified diagrammatic representations of the coefficients
$A_6,A_4,A_2$, we now need to simplify all diagrams to reduce their
computational complexity, as discussed in Section \ref{sec:diagrams}.
Eliminating all filled dots in regions that address only one tensor
yields:
\begin{align}
  A_6 = D_1 - D_2 + D_3 - D_4,
\end{align}
\begin{align}
  A_4 = D_2 - 2 D_3 + 3 D_4,
  &&
  A_2 = D_3 - 3 D_4
\end{align}
where
\begin{align}
  D_1 \equiv \f18 \diagram{triple_0_o6},
  &&
  D_2 \equiv \f32 \diagram{triple_01_o4}
  - 2 \diagram{triple_1_o3},
  &&
  D_4 \equiv \diagram{triple_0111}.
\end{align}
\begin{align}
  D_3 \equiv 3 \diagram{triple_011_o2}
  + \f34 \diagram{triple_02_o2}
  - 6 \diagram{triple_11_o}
  + 2 \diagram{triple_2}.
\end{align}
We can also write the product $\X$ in \eqref{eq:triple_multi} using
collective operators as
\begin{align}
  \X \EQFS
  \tilde A_6 \col{X}^6 + \tilde A_4 \col{X}^4 + \tilde A_2 \col{X}^2,
  \label{eq:triple_col}
\end{align}
where we have neglected scalar terms without physical consequence, and
\begin{align}
  \tilde A_6 \equiv \f{A_6}{N!_6},
  &&
  \tilde A_4 \equiv \f{A_4}{N!_4} - 5\p{3N-8} \f{A_6}{N!_6},
\end{align}
\begin{align}
  \tilde A_2 \equiv \f{A_2}{N!_2} - 2\p{3N-4} \f{A_4}{N!_4}
  + \p{45N^2-210N+184} \f{A_6}{N!_6}.
\end{align}
